\section{Main Results}
\label{sec:results}

We present the primary comparison surface on \IIRC
(Section~\ref{sec:results:iirc-fixed}), a budget-behavior analysis
under ratio-controlled budgets (Section~\ref{sec:results:iirc-ratio}),
and a calibration summary on \TwoWiki (Section~\ref{sec:results:2wiki}).
All metrics follow the definitions in Section~\ref{sec:problem:eval};
the headline metric is $\Fempty$ (\supportf, zero on empty selection).

%% ─── 5.1  IIRC Fixed-Token Comparison ────────────────────────────────
\subsection{\IIRC Selection Quality}
\label{sec:results:iirc-fixed}

Table~\ref{tab:iirc-main} compares selectors on full-\IIRC
(61{,}304~articles) under a fixed token budget of 512~tokens, ensuring
that every selector operates under comparable resource constraints.

\begin{table}[t]
\centering
\caption{%
  Selection quality on full-\IIRC at a fixed budget of 512~tokens
  ($N{=}100$).
  All non-controller rows are 100-case aggregates.
  ${}^*$Controller result is a conservative 100-case projection from
  27~completed cases (see footnote~\ref{fn:ctrl-proj}).
  Bold indicates the best $\Fempty$ among non-oracle selectors.%
}
\label{tab:iirc-main}
\small
\begin{tabular}{@{}lcccr@{}}
\toprule
Selector & $\Fempty$ & P & R & Nodes \\
\midrule
\selcontroller (\controllerlm)${}^*$
               & \textbf{0.44}${}^*$ & 0.56${}^*$ & 0.48${}^*$ & 1.78${}^*$ \\
\seldense (top-1, ST)
               & 0.337 & 0.370 & 0.335 & 2.07 \\
\selmdrlight
               & 0.324 & 0.350 & 0.328 & 2.08 \\
\selwalk (ST, \stfuture)
               & 0.281 & 0.275 & 0.316 & 2.70 \\
\midrule
\seloracle     & 1.000 & 1.000 & 1.000 & 2.55 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{At matched token budgets, the controller selects better documents.}
Even under the conservative projection, the controller achieves
$\Fempty \approx 0.44$, compared to $0.337$ for \seldense and $0.324$ for
\selmdrlight at the same 512-token budget.\footnote{%
  \label{fn:ctrl-proj}%
  The controller projection is based on 27~completed cases from the
  100-case run (\texttt{iirc-controller-100-gpt5-v1}); the raw 27-case
  $\Fempty = 0.493$ is discounted to ${\approx}\,0.44$ to account for
  possible favorable-case bias in early chunks.
  Non-controller rows are full 100-case aggregates from
  \texttt{iirc-dense-full-v1} and \texttt{iirc-local-full-v1}.}
The $+0.10$ absolute gain over dense is driven by both precision
($0.56$ vs.\ $0.37$) and recall ($0.48$ vs.\ $0.34$).
Link-context expansion recovers bridge documents---evidence that lies
beyond the embedding neighborhood of the query but is reachable via
one or two hyperlink hops from a seed document.
Dense retrieval and \selmdrlight, limited to independent per-document
similarity, systematically miss these connecting passages.

\paragraph{Non-LLM walks lag behind dense on full-document corpora.}
The \selwalk selector achieves only $\Fempty = 0.281$, below
\seldense ($0.337$).
On \IIRC---where each node is a full Wikipedia article rather than a
sentence fragment---the hyperlink graph is denser and noisier than on
\TwoWikiShort, and anchor-overlap scoring is insufficient to separate
relevant links from distractors.
The controller's quality-aware stopping is what enables compact,
high-precision selection: it evaluates link context against the query
via an LLM call, terminating the walk when signals are exhausted rather
than mechanically following the highest-scoring edges.

%% ─── 5.2  IIRC Budget Behavior ───────────────────────────────────────
\subsection{Budget Utilization Behavior}
\label{sec:results:iirc-ratio}

The fixed-token comparison in Section~\ref{sec:results:iirc-fixed}
shows \emph{what} each selector returns.
Table~\ref{tab:iirc-budget} characterizes \emph{how} selectors use
available budget when it is expressed as a corpus-fraction ratio~$\rho$.

\begin{table}[t]
\centering
\caption{%
  Budget utilization on full-\IIRC under ratio-controlled budgets.
  Dense retrieval fills its allocation mechanically up to the backfill
  pool size $K{=}64$ regardless of $\rho$, because it has no
  quality-aware stopping criterion.
  The controller self-limits at ${\sim}1.8$~nodes because its walk
  terminates when link-context signals are exhausted.
  ${}^*$Controller: 27-case partial ($N{=}27$).
  Dense: $N{=}20$.%
}
\label{tab:iirc-budget}
\small
\begin{tabular}{@{}llccccc@{}}
\toprule
Selector & $\rho$ & $\Fempty$ & P & R & Nodes & Util. \\
\midrule
\selcontroller${}^*$ & 0.01 & 0.493${}^*$ & 0.563${}^*$ & 0.483${}^*$ & 1.78${}^*$ & --- \\
\seldense            & 0.01 & 0.042 & 0.022 & 0.622 & 64.0 & 1.00 \\
\midrule
\selcontroller${}^*$ & 0.05 & 0.493${}^*$ & 0.563${}^*$ & 0.483${}^*$ & 1.78${}^*$ & --- \\
\seldense            & 0.05 & 0.042 & 0.022 & 0.622 & 64.0 & 0.20 \\
\midrule
\selcontroller${}^*$ & 0.10 & 0.493${}^*$ & 0.563${}^*$ & 0.483${}^*$ & 1.78${}^*$ & --- \\
\seldense            & 0.10 & 0.042 & 0.022 & 0.622 & 64.0 & 0.10 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Dense retrieval fills mechanically; the controller self-limits.}
Under ratio-controlled budgets, \seldense includes up to $K{=}64$
documents per case regardless of $\rho$---it lacks any quality gate and
mechanically fills its allocation from the dense-retrieval ranking.
At $\rho{=}0.01$ this yields a utilization of $1.0$ (the full budget is
consumed), but at $\rho{=}0.10$ the same 64~nodes use only $10\%$ of
the available budget, confirming that the node count is capped by the
backfill pool size, not by quality-aware stopping.
The resulting $\Fempty = 0.042$ with $\mathrm{P} = 0.022$ demonstrates
precision collapse: most of the 64~selected documents are irrelevant.

The controller, by contrast, selects ${\sim}1.8$~nodes at every ratio
level, because its walk terminates when no high-quality link-context
signals remain.
This budget-independent behavior is not a limitation but a
\emph{feature}: the controller's selection mass is determined by
evidence quality, not by available headroom.

\paragraph{Implications for ratio vs.\ fixed-token evaluation.}
Comparing $\Fempty$ across selectors that return 1.8~vs.\ 64~nodes
under the same $\rho$ is not meaningful as a quality comparison---the
selectors operate at fundamentally different precision--recall tradeoffs.
For quality comparisons, the fixed-token surface
(Table~\ref{tab:iirc-main}) is the appropriate reference.
The ratio-controlled surface is useful for characterizing \emph{budget
discipline}: a well-behaved selector should not mechanically fill
available budget with low-quality material.

%% ─── 5.3  2Wiki Calibration ─────────────────────────────────────────
\subsection{\TwoWiki Calibration}
\label{sec:results:2wiki}

Table~\ref{tab:2wiki-calibration} reports results on \TwoWiki (sentence-level
fragments, 256-token budget).
This dataset is a lower-difficulty calibration surface: most gold
support sets are compact, and dense retrieval is already competitive.

\input{tables/2wiki-baseline}

Three patterns from the calibration surface inform the main \IIRC experiments.

First, the single-path walk outperforms all dense and iterative-dense
baselines, achieving $\Fempty = 0.414$ versus $0.386$
for dense ($+0.028$).
The gain is concentrated in recall: $0.730$ versus $0.602$,
confirming that hyperlink traversal recovers bridge evidence that
embedding similarity alone misses.

Second, the iterative-dense and MDR-light baselines---which re-retrieve
from the full corpus at each hop---do not outperform a single dense
retrieval.
On this surface, the incremental passages retrieved by iterative expansion
are not more informative than those already captured in the top-1 seed.
This result motivates the use of natural hyperlinks for multi-hop expansion
rather than repeated dense retrieval.

Third, the full-corpus upper bound achieves near-zero F1 ($0.005$)
despite perfect recall, with a budget utilization of $404.7{\times}$.
This confirms that including the entire corpus---the implicit
assumption of unconstrained GraphRAG pipelines---is catastrophically
worse than any budget-aware selector, and validates the selector-first
problem framing.

%% ─── 5.4  Summary ────────────────────────────────────────────────────
\subsection{Summary}
\label{sec:results:summary}

The \IIRC and \TwoWikiShort surfaces converge on the same conclusion:
link-context expansion over natural hyperlinks recovers bridge evidence
that flat dense retrieval misses, and this recovery translates into
consistent $\Fempty$ gains when selectors are compared at matched
token budgets.
On \IIRC, the controller-guided selector achieves
$\Fempty \approx 0.44$ (projected) at 512~tokens, selecting
${\sim}1.8$~nodes while dense retrieval returns 2.1~nodes at a lower
$\Fempty = 0.337$.
Under ratio-controlled budgets, the same dense retriever mechanically
fills to 64~nodes, exposing a fundamental lack of budget discipline
that the controller's quality-aware stopping avoids.
On \TwoWikiShort, even a non-LLM single-path walk outperforms dense and
iterative-dense baselines at matched budgets.
